Analyzing the Effectiveness of Copyright Law on AI Output in Software Engineering and Future Developments

### 2025-02-18 paper/main.tex
mostly done with paper

\textbf {
  For those concerned with the legality of the software they produce, the usage of
AI-generated code should be avoided as their trustworthiness is dubious and
they will become succeptible to copyright law in the near future.
}

@@

@@

% \\
%thesis

@@

@@

% TODO: maybe change this name?
\section{Tracking}

@@

@@

% use section* for acknowledgment
\section*{Acknowledgment}
The authors would like to thank...

>>>>

\textbf {
  For those concerned with the legality of the software they produce, the usage of
AI-generated code should be avoided as soon as possible, as they could become succeptible to
copyright law in the near future.
}

@@

Copyright is a mechanism used to protect created works from direct copying
but not a sufficiently derivative use or one that gives adequate credit to
the original author \cite[p.~55]{stokes2021}. There has been one landmark
case regarding copyright and programming, Google INC v. Oracle America, Inc,
which regards the copying of generic and widely known API code by Google
from Oracle \cite{harvard}. API in short means that its code that acts as a
abstracted layer, such that the user doesnt have to think about whats going
on underneath. Think of it like me making my own laptop, but still keeping
around the QWERTY keyboard so the users of my laptop dont have to learn a
whole new keyboard layout. The ruling stated that its alright for google to
copy this code because they were doing it in the interest of the general
public and thought it wasnt possible to copyright something so generic,
leading to google winning the case.

@@

%thesis

@@

AI has been getting better and better at generating
code \cite{codesignal}, due to the vasty increasing training data
these models are harvesting from the web \cite{lacour_2024}.
Especially with many popular code repositories being open online,
its fairly straightforward to assume that they are being used for training.
While its impossible to know exactly whats being used as training data and
what isnt by these large companies, they arent making it any easier with their
lack of clarity and reassurance \cite{willison_2023}. Therefore, it can be
reasonably assumed that licensed code that is public purview is also
being used as training data.

The issue then arises with the licensure of the code ingested by the model.
Whether its under a "Proprietary" license or a "Copy Left" license, the
end user could be in trouble.

A Proprietary license would bar any derivative usage, putting the
user in deep trouble if found out, whether they're trying to publicize
their code or not.

A Copy Left, while less troublesome, an entity trying to privatize their code
would fail to uphold the conditions of the license, which demands that any derivative
works also be in the public domain.

We havent even gotten to the fact that many licenses require attribution to the
original author, which would be completely lost in this process of ai ingestion
and generation.

@@

% TODO: maybe change this name?
\section{Detecting the Origins of Generated Code}

While it may seem impossible, If there was a way to figure out where
the code generated by an AI came from, any users of the generated code
could be in huge trouble depending on the licensure of the source code.
Such technology doesnt exist today but is tending towards that direction,
as seen in this study: \cite{ma2024}. They dive deep into the possibility
of detecting whether or not code generated by a model can be traced
back / verified to be in some dataset.

Obviously this hasn't been attempted on mainstream models, but
with their findings being so fruitful, its only a matter of time
before new research builds off this foundation to see if
generated code has its roots on public code. Logically then,
source code authors could audit whether or not others have indirectly used their
source code, via this AI ingestion and generation proxy. Now a solid case
can be built up by authors with strict licenses, that this usage is
in violation of their licenses and the derivate users now face
legal trouble.

\section{Forward Compliance}

While origin analysis of AI-generated code is not yet developed, preventative
action must be taken now by those interested in upholding the legality of their
software. As this technology develops, the risk of legal trouble only increases.

Forward compliance simply means to disassociate with AI-generation tools, and
return to a "traditional" style of programming. This is doubley-so for corporate
projects which would face even higher repurcussions by any future lawsuits.

While some may argue that the process of AI-generating code is transformative by
defenition, but currently there exists no precedent, so claming as such is meaningless.
When dealing with Copyright, its best to live on the side of pessimism, so
treating AI-generation as akin to copying allows users of these tools
to be protected in case future precedent is pessimistic as well.

Others can argue that the copying of code has precedent, and is allowed according
to Google INC v. Oracle America Inc., but a deeper inspection of the case details must
be had. The copying of the API code was only allowed because its value lay
in the familiarity the general public already had with it. Additionally this API code
lacked any of the deeper functionality which could be copyrighted and protected, its sematics
ruled to be too abstract \cite{harvard}.

This reasoning cannot be used for all code however,
as the AI-tools can generate anything from this API-like code, to code that performs
real work, with no care for the licensure of either. If an author were to form a case
claiming that their proprietary code was "copied" and now exists with an unauthorized
holder, using the afforementioned origin anazlyzer to show the means of "copying,
the results of this case would turn out very different.

@@

There exists a real legal threat, maybe not now but certainly in the near future,
for those using AI-generated code within their produts, doubly-so if its commercial
in nature. As of right now there is no way to tell if the code generated comes
from a open-source repository or one thats heavily licensed. With the research
on origin analyzers only developing, its very reasonable for AI-generated code
to be back tracked to its source, leading users to fall under jurisdiction of
this source code. For those concerned with the legality of their code, its
best to stray away from these tools.

@@

% use section* for acknowledgment
\section*{Acknowledgments and Notes}
I'd like to thank Dr.Kirsch and my peer reviewers for helping this essay come together. 
One thing I'd like to note are that the text formatter im using for this template does
not allow for citations to be bolded, so note that any references to \cite{stokes2021},
\cite{harvard} and \cite{ma2024} are my peer-reviewed sources, sorry for the trouble.
